---
output: md_document
---
# How to get move ticks closer to the axis?

To answer this question let's start off by loading some packages and preparing userRs favorite dataset `mtcars` for plotting.

```{r message=FALSE}
library(dplyr)
library(ggplot2)
data("mtcars")

cyl <- count(mtcars, cyl) %>%
  arrange(n) %>%
  mutate(cyl = reorder(cyl, n))

cyl
```

Next, let's create a basic plot.

```{r}
p1 <- ggplot(cyl, aes(cyl, n)) +
  geom_col() +
  theme_minimal() +
  coord_flip()
p1
```

That's already quite nice but for a bar chart the horizontal grid lines are redundant. Let's remove them.

```{r}
p2 <- p1 + theme(
  panel.grid.major.y = element_blank(),
  panel.grid.minor.y = element_blank()
)
p2
```

Great! But now the axis ticks are quite far from the bars. Let's move them closer.

```{r}
p3 <- p2 + theme(axis.text.y = element_text(margin = margin(r = -15)))
p3
```

And there you have it! That's how you move axis ticks closer to the axis. The `r` argument within `margin()` stands for right. Accordingly, there are also `l`, `t` and `b` for left, top and bottom, respectively.

Next, let's try to move the x axis ticks up. Notice the `axis.text.x` instead of `axis.text.y`.

```{r}
p3 + theme(axis.text.x = element_text(margin = margin(t = 15)))
```

Huh, that was unexpected: the ticks went down instead of up. Let's try passing a negative number.

```{r}
p3 + theme(axis.text.x = element_text(margin = margin(t = -15)))
```

That did the job! It's probably a bit too high but you get the idea.
